Analysing Constraint Grammars with a SAT-solver

### 2016-03-08 paper/lrec/eval_futurework.tex
trying to explain the podar/sentar example nicer; maybe I should just remove it?

\subsection{Dutch} The Dutch grammar had two kinds of errors: rule-internal, and rule interaction. As for rule-internal conflicts, one was due to a misspelling in the list definition for personal pronouns, which rendered 5 rules ineffective. The other was about subreadings: the genitive \emph{s} is analysed as a subreading in the Apertium morphological analyser, but it appeared in 2 rules as the main reading.

@@

These two rules both remove an adverb, but the first one has a broader condition:
remove adverb if followed by a noun. In contrast, the second rule has a stricter condition: only remove adverb if it is preceded by a determiner, the adverb itself is ambiguous with an adjective, and followed by a noun. The problem is that the first rule removes the adverb in all possible cases, and the second will not have any chance to act. If the rules were in the opposite order, then there would be no problem.

@@

As for the 25 interaction conflicts, there were only 9 distinct rules that rendered 25 other rules ineffective.
In fact, we can reduce these 9 rules further into 3 different groups: 4 + 4 + 1, where the groups of 4 rules are variants of otherwise identical rule, with  combinations feminine singular, masculine singular, feminine plural and masculine plular.
An example of such conflict is below (gender and number omitted for readability):

@@

For a full list of found conflicts, see \todo{\url{actual://link.here/spanish/error.log}}

@@

The comments make it clear that the first rule is meant to disambiguate between \emph{poder} and \emph{podar}, and the second rule between \emph{sentir} and \emph{sentar}. But the rule does not mention anything about \emph{poder} nor \emph{sentir}; without the ambiguity class constraints, the SAT-solver could give a triggering sequence with any other reading in the target.

This example shows the 
shows a shortcoming in our way of handling the lexical conditions. 
If the lemmas and word forms were integrated to the analyses in a more
sensible way, then the ambiguity class constraints would create a
sequence where \emph{podar} was ambiguous with \emph{poder} only in the
word form \emph{puedo}.

instead of randomly chosen readings. So, even though this was not marked as a conflict, we can say that the weird output of the tool was unnecessary, and by better implementation, the tool would have created more realistic example sequence.

>>>>

\subsection{Dutch} The Dutch grammar had two kinds of errors: rule-internal, and rule interaction. As for rule-internal conflicts, one was due to a misspelling in the list definition for personal pronouns, which rendered 5 rules ineffective. The other was about subreadings: the genitive \emph{s} is analysed as a subreading in the Apertium morphological analyser, but it appeared in one rule as the main reading.

@@

% These two rules both remove an adverb, but the first one has a broader condition:
% remove adverb if followed by a noun. In contrast, the second rule has a stricter condition: only remove adverb if it is preceded by a determiner, the adverb itself is ambiguous with an adjective, and followed by a noun. 
The problem is that that
The first rule has a broader condition, 
it removes the adverb in all possible cases 
 removes the adverb in all possible cases, and the second will not have any chance to act. If the rules were in the opposite order, then there would be no problem.

@@

As for the 25 interaction conflicts, there were only 9 distinct rules that rendered 25 other rules ineffective.
In fact, we can reduce these 9 rules further into 3 different groups: 4 + 4 + 1, where the groups of 4 rules are variants of otherwise identical rule, each with different gender and number.
An example of such conflict is below (gender and number omitted for readability):

@@

For a full list of found conflicts, see error log\footnote{\url{https://github.com/inariksit/cgsat/raw/master/data/spa/conflicts.log}}.

@@

The comments make it clear that the first rule is meant to disambiguate between \emph{poder} and \emph{podar}, and the second rule between \emph{sentir} and \emph{sentar}. But the rule does not mention anything about \emph{poder} nor \emph{sentir}.
Even with the ambiguity class constraints, the SAT-solver only created an ambiguity where \emph{sentar} in 1st person plural is ambiguous with an anonymous 1st person plural reading.
This does not reflect the reality, where these targets are only ambiguous with certain verbs, and in certain conjugated forms.

Another question is whether this level of detail is necessary. 
After all, the grammar will be used to disambiguate real life texts, where \emph{podar} will likely not have any other ambiguities. 
If our handling of lexical conditions was better, this sequence would, in fact, only be ambiguous with \emph{poder} in the right conjugation form.

% shows a shortcoming in our way of handling the lexical conditions. 
% If the lemmas and word forms were integrated to the analyses in a more
% sensible way, then the ambiguity class constraints would create a
% sequence where \emph{podar} was ambiguous with \emph{poder} only in the
% word form \emph{puedo}.

% instead of randomly chosen readings. So, even though this was not marked as a conflict, we can say that the weird output of the tool was unnecessary, and by better implementation, the tool would have created more realistic example sequence.
